---
title: "Exam 1"
author: "Isabella Werneck"
date: "10/04/2021"
output: html_document
---


# Instructions

a. Create a folder in your computer (a good place would be under Crim 250, Exams). 

b. Download the dataset from the Canvas website (fatal-police-shootings-data.csv) onto that folder, and save your Exam 1.Rmd file in the same folder.

c. Download the README.md file. This is the codebook. 

d. Load the data into an R data frame.

```{r}

dat <- read.csv(file = 'fatal-police-shootings-data.csv')

```


# Problem 1 (10 points)

a. Describe the dataset. This is the source: https://github.com/washingtonpost/data-police-shootings . Write two sentences (max.) about this.


__Answer: The dataset contains data on every fatal shooting in the United States by a police officer in the line of duty since Jan. 1st, 2015. The information was collected and compiled by The Washington Post, and it contains information on the names, races, and ages of the victims, the circumstances of the deaths, and the locations of the shootings.__


b. How many observations are there in the data frame?
```{r}
dim(dat)
```

__Answer: There are 6594 observations in the dataset.__


c. Look at the names of the variables in the data frame. Describe what "body_camera", "flee", and "armed" represent, according to the codebook. Again, only write one sentence (max) per variable.

```{r}
names(dat)
```

__Answer:__

__`body_camera` represents whether news reports have indicated that the armed officer was wearing a body camera during the incident.__

__`flee` represents whether news reports have indicated that the victim was moving away from the officers during the incident.__

__`armed` represents whether the victim was armed with an object that a police officer believes could inflict harm.__



d. What are three weapons that you are surprised to find in the "armed" variable? Make a table of the values in "armed" to see the options.
```{r}

table(dat$armed)

```

__Answer: I was surprised to see "air conditioner", "stapler", and "microphone".__ 


# Problem 2 (10 points)

a. Describe the age distribution of the sample. Is this what you would expect to see?
```{r}

#Histogram for age
hist(dat$age, main = "Ages in Fatal Police Shootings Dataset", xlab = "Age", ylab = "Frequency")

```

__Answer: The histogram of the dataset shows that the most frequent age in the dataset is 30-35 years old. The great majority of the individuals are between ages 20-40. This is what I would expect to see in this study, since the majority of criminals, as well as people targeted by the police, are adults and young adults.__


b. To understand the center of the age distribution, would you use a mean or a median, and why? Find the one you picked.
```{r}


summary(dat$age)

```

__Answer: I would use median because it is more representative of the center of age distribution. It is more resistant to errors (for example, older people writing down the wrong age), and it is not influenced by high or low observations, like 91 or 6 years old. Especially in a large dataset like this one, the median is more representative of the age at the center.__

c. Describe the gender distribution of the sample. Do you find this surprising?


```{r}
table(dat$gender)
```


```{r}
counts <- table(dat$gender)
barplot(counts, main = "Gender in Fatal Police Shootings Dataset", xlab = "Gender", ylab = "Frequency" )

```

__Answer: The barplot and the table above show that there is a significantly higher number of males than females who have been killed by an armed officer in the United States since 2015. The first bar contains 3 observations with missing values. The number is low enough in comparison with the other datapoints (3 vs 293 and 6298) that we can disregard it in the analysis of this data. I did not find the gender distribution surprising — there are more male criminals than female criminals, and males are more likely to be seen as a threat to police officers than females. This is due to many reasons, including the disparities in upbringing between males and females, as well as the societal expectations that vary between genders.__



# Problem 3 (10 points)

a. How many police officers had a body camera, according to news reports? What proportion is this of all the incidents in the data? Are you surprised that it is so high or low?

```{r}
table(dat$body_camera)

```

__Answer: In the table above, `False` means that the police officer did not wear a body camera at the time of the incident, and `True` means that they did. Out of 6594 officers, 910 officers wore body cameras, and 5684 officers did not. I am surprised that only 910 wore body cameras, since I thought that police officers were obligated to wear them. However, only a few states currently require the use of body cameras, which makes sense with the results. __


b. In  how many of the incidents was the victim fleeing? What proportion is this of the total number of incidents in the data? Is this what you would expect?
```{r}

table(dat$flee)

```


__Answer: In the table above, `Car` means the victims were fleeing by car, `Not fleeing` means that the victim did not attempt to flee, and `Foot` means that the victims were fleeing by foot. The unlabeled 491 are missing data; since this dataset was collected from news reports, the amount of missing data is normal. Reporters might not have made it to the scene in time, and people involved might not have been willing to talk about the incident. This should be taken into consideration in analyzing the quality of the data, but these numbers can be disregarded for the purposes of this question. `Other`, which represents 248 cases, can be interpreted in several different ways. It could mean that the victim attempted to flee both by foot and by car, and it could also mean that they fled in another vehicle, like trains or buses. I chose to interpret it as attempts to flee that do not fall into the categories `Car` or `Foot` since the cases do not fall into `Not fleeing`. If we add Other, Foot, and Car, we get 2151 "Fleeing" cases, which represents a smaller group of cases than `Not fleeing`. This is exactly what I expected, since I imagine that most individuals would avoid fleeing from an armed officer.__



# Problem 4 (10 points) -  Answer only one of these (a or b).

a. Describe the relationship between the variables "body camera" and "flee" using a stacked barplot. What can you conclude from this relationship? 

*Hint 1: The categories along the x-axis are the options for "flee", each bar contains information about whether the police officer had a body camera (vertically), and the height along the y-axis shows the frequency of that category).*

*Hint 2: Also, if you are unsure about the syntax for barplot, run ?barplot in R and see some examples at the bottom of the documentation. This is usually a good way to look up the syntax of R code. You can also Google it.*


```{r}

tab.fleebodycamera <- table(dat$body_camera, dat$flee)
barplot(tab.fleebodycamera, ylim=c(0, 4000),
        main = "Flee vs Body Camera in Fatal Police Shootings Dataset",
        xlab = "Victims that Attempted to Flee/Not Flee", ylab = "Frequency of Cases",
        legend.text = rownames(tab.fleebodycamera),
        beside = FALSE) # Stacked bars (default))
```


__Answer: I can conclude that there is a proportional number of officers that wore body cameras in each category of "Fleeing". The barplot does not show a strong correlation between whether the officers wore body cameras and the victims' attempts to flee.__ 


b. Describe the relationship between age and race by using a boxplot. What can you conclude from this relationship? 

*Hint 1: The categories along the x-axis are the race categories and the height along the y-axis is age.* 

*Hint 2: Also, if you are unsure about the syntax for boxplot, run ?boxplot in R and see some examples at the bottom of the documentation. This is usually a good way to look up the syntax of R code. You can also Google it.*


```{r}

```

__Your answer here.__






# Extra credit (10 points)

a. What does this code tell us? 

```{r, eval=FALSE}
mydates <- as.Date(dat$date)
head(mydates)
(mydates[length(mydates)] - mydates[1])
```

__Answer: The code shows that the dataset was collected between 2458 days. This means that the between the first case and the last case, 2458 days have passed.__ 

b. On Friday, a new report was published that was described as follows by The Guardian: "More than half of US police killings are mislabelled or not reported, study finds." Without reading this article now (due to limited time), why do you think police killings might be mislabelled or underreported?

__Answer: Police killings might be mislabelled or underreported because not a lot of officers utilize body cameras, which would show precisely what occured during the incidents, and because a lot of the data is collected by journalists/reporters. It is difficult for them to be at the scene to collect data, and they might not have access to police reports due to sensitive information. Another reason is because many of these cases might contain information that could end police officers' careers and possibly even put that in prison. There are many cases of police brutality and murder in the US, and many go without any coverage.__ 

c. Regarding missing values in problem 4, do you see any? If so, do you think that's all that's missing from the data?



